---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covid19br

<!-- badges: start -->
<!-- badges: end -->

The goal of the package is to provide a function to automatically import the Brazilian COVID-19 pandemic data into R. Brazilian data is available on country, region, state, and city levels, and are downloaded from the official Brazilian repository at &lt;https://covid.saude.gov.br/>. The package also enables one to download the world-level COVID-19 data from the Johns Hopkins University's repository at &lt;https://github.com/CSSEGISandData/COVID-19>.

## Installation

You can install the released version of covid19br from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("covid19br")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fndemarqui/covid19br")
```
## Example 

Downloading COVID-19 pandemic data:

```{r example, message = FALSE}
library(covid19br)
library(tidyverse)

brazil <- downloadCovid19("brazil")
regions <- downloadCovid19("regions")
states <- downloadCovid19("states")
cities <- downloadCovid19("cities")
world <- downloadCovid19("world")

```

