---
title: "Introduction to the covid19br package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the covid19br package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette shows how to use the **covid19br** R package for downloading and exploring data from the COVID-19 pandemic in Brazil and the globe as well. The package downloads data bases from the following repositories:

- The official brazilian repository provided by the brazilian government: [https://covid.saude.gov.br](https://covid.saude.gov.br); 

- The Johns Hopkins University's repository: [https://github.com/CSSEGISandData/COVID-19](https://github.com/CSSEGISandData/COVID-19) 

The last repository has data on the COVID-19 pandemic in the global level (daily counts of confirmed cases, deaths and recovered patients by countries and territories), and has been widely used all over the world as a reliable source of data information on the COVID-19 pandemic. The former repository, on the other hand, possesses data on the Brazilian territory by city, state, region and national levels. 

We hope that this package may be helpful to other researchers and scientists to understand and fight this terrible epidemic that has been plaguing the world.

# Package installation

The development version of the  R package ***covid19br*** can be found at [github.com/fndemarqui/covid19br](https://github.com/fndemarqui/covid19br), and it can be installed using the following command line in R: devtools::install_github("fndeamrqui/covid19br"). 

The installation procedure described abouve requires the package ***devtools*** to be previously installed in R.

# Getting started with covid19br R package

We will get started by showing how to use the package to load into R data sets of the COVID-19 pandemic by downloading the COVID-19 data set from the official Brazilian repository [https://covid.saude.gov.br](https://covid.saude.gov.br)


```{r, message = FALSE}
library(covid19br)
library(tidyverse)

brazil <- downloadCovid19("brazil")

ggplot(brazil, aes(x = date, y = accumDeaths)) +
  geom_point() +
  geom_path()


```
